---
title: "Engagement with population health care transition tool among those with limited English proficiency (ver 4)"
date: "2019-11-24"
output: pdf_document
author: "Margaret Wheeler, Kristin Gagliardi, Lev Malevanchik, Sachin Shah"
#classoption: landscape
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)    # plotting data
library(tableone)
library(haven)

knitr::opts_chunk$set(echo = FALSE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width=6, fig.height=4)  
knitr::opts_chunk$set(comment = NA)
options(ztable.type = 'html')

#Custom color palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for line and point colors, add
scale_colour_manual(values=cbbPalette)

## FIGURE SETTINGS
sachin_theme <- theme_bw()+
        theme(
          panel.background = element_blank(), 
          axis.line = element_line(colour = "grey75"),
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey", linetype = "dotted" ),
          panel.grid.minor.y = element_blank(),
          legend.position = "none", 
          axis.title=element_text(size=10)
          )

sachin_theme_leg <- theme_bw()+
        theme(
          panel.background = element_blank(), 
          axis.line = element_line(colour = "grey75"),
          panel.border = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(color = "grey", linetype = "dotted" ),
          panel.grid.minor.y = element_blank(),
          legend.position = "top", 
          legend.title = element_blank(),
          axis.title=element_text(size=10)
          )



```


###4.1 Patient characteristics by LEP
* Patients characterized as limited English prociency (LEP) if they are noted to require an interpreter or list a preferred language other than English
* Patient characteristics pulled from APeX 

###Table 1: Patient characteristics by English proficiency
```{r table1, echo=FALSE}

### LOAD DATA ###
t1 <- read_sas("processed data/ctop_apex_first_20200605.sas7bdat", 
    NULL)

myVars <- c("age", "sex", "recode_race", "Ethnicity", "Marital_Status", "ins", "apex_lang_cat", "INTRPTR_NEEDED_YN", "Top3_service","LOS","ICU", "Elixhauser", "DRG_Wt", "Disposition")

catVars <- c("sex", "recode_race", "Ethnicity", "Marital_Status", "ins", "apex_lang_cat", "INTRPTR_NEEDED_YN", "Top3_service","ICU", "Disposition")

not_nl <- c("age", "LOS","DRG_Wt")

library(labelled)
var_label(t1) <- list(age = "Age", sex = "Male", recode_race = "Race", Ethnicity = "Hispanic ethnicity", 
                     Marital_Status = "Marital status", ins = "Insurance status",
                     apex_lang_cat = "Language", INTRPTR_NEEDED_YN = "Interpreter used", 
                     Top3_service = "Clinical service", LOS = "Length of stay", 
                     ICU = "ICU stay", Elixhauser = "Elixhauser score",
                     DRG_Wt = "DRG weight", Disposition = "Discharge home with services")


tab2 <- CreateTableOne(vars = myVars, strata = "EP", data = t1, factorVars = catVars)
tab2a <- print(tab2, nonnormal = not_nl, catDigits = 0, contDigits = 1, varLabels = TRUE, format = "fp", test = T, noSpace = TRUE, dropEqual = TRUE, printToggle = TRUE, quote = T)
 # tab2b <- tab2a[,1:3]
 # print(tab2b, quote = FALSE)

```




###4.4 What is the association between LEP and post-discharge issues

### Figure 2: Identified issue by English proficiency
```{r 44, include = TRUE, fig.height=4, fig.width=8}

library(dplyr)
library(scales)
library(tidyverse)
#### LOAD DATA ####
library(readr)
issue_rates <- read_csv("C:/Users/sachi/Box Sync/CTOP/Results for paper/CTOP LEP issue rates 2020-06-05.csv")
issue_rates <- issue_rates %>% mutate(issue=recode(issue, 
                        'Discharge instruction'='Discharge instructions',
                        'Medications' = 'Medication concerns',
                        'Prescriptions' = 'Getting prescriptions',
                        'Follow up help' = 'Follow up questions',
                        'Symptoms' = 'New or worsening symptoms',
                        'Other clinical issues' = 'Any other clinical issues'))
issue_rates$issue <- as.factor(issue_rates$issue)
issue_rates$issue <- factor(issue_rates$issue, levels = c("Discharge instructions", "Getting prescriptions", "Medication concerns", "Follow up questions", "New or worsening symptoms", "Any other clinical issues"))
issue_rates$EP <- factor(issue_rates$EP, levels = c("LEP", "EP"))



## ADJUSTED- BAR ##
issue_rate_fin_adjusted <- issue_rates  %>% 
                            filter(adjusted == "Adjusted") %>%
                            mutate(sig = "Y") %>%
                            mutate(aster = ifelse(sig == "Y" & EP == "LEP","*",NA))

aster <- issue_rate_fin_adjusted %>%
            filter(aster == "*") %>%
            select(issue, freq, aster, EP) %>%
            mutate(freq = freq + 0.01)


pd <- position_dodge(0.2)
fig_issue_rate_adj_bar <- ggplot(data = issue_rate_fin_adjusted, aes (x= issue, y = freq, fill = EP)) +
          geom_bar(stat = "identity", position = "dodge", width = .5) +
          scale_y_continuous(name = "Predicted likelihood", 
                             limits = c(0.0,0.2), breaks = c(seq(from = 0.0, to = 0.20, by = 0.05)),
                             labels = scales::percent_format(accuracy = 1)
                             ) +
          scale_x_discrete(name = "Post-discharge issues", labels = wrap_format(10)) +
          geom_text(data = aster, label = "*", size = 7)+ 
          sachin_theme_leg

fig_issue_rate_adj_bar
#ggsave("C:/Users/sachi/Box Sync/CTOP/Results for paper/fig_issue_rate_adj bar 2020-10-12.png", height = 4, width = 6.5)
ggsave("C:/Users/sachi/Box Sync/CTOP/Results for paper/fig_issue_rate_adj bar 2020-10-12.tiff", height = 4, width = 6.5)



```



### Figure 3: Involvement of other health professionals among those who identified an issue by English proficiency
```{r involved, fig.width=6, fig.height=4}

library(dplyr)
library(scales)

#### LOAD DATA ####
library(readr)
involved <- read_csv("C:/Users/sachi/Box Sync/CTOP/Results for paper/CTOP clinician involvement rates 2020-06-05.csv")
involved$involved <- as.factor(involved$involved)
involved$EP <- factor(involved$EP, levels = c("LEP", "EP"))

#### UNADJUSTED ####
involved_un <- involved  %>% filter( adjusted == "Unadjusted")

fig3_un <- ggplot(data = involved_un, mapping = aes (x= involved, y = freq)) +
          geom_point(aes(color = EP), size = 3 , position = pd) +
          geom_errorbar(aes(ymin = LL, ymax = UL, width = 0.1, color = EP), position = pd) +
           scale_y_continuous(name = "Percent with involvement", 
                             limits = c(0.0,0.11), breaks = c(seq(from = 0.0, to = 0.10, by = 0.02)),
                             labels = scales::percent_format(accuracy = 1)
                             ) +
          scale_x_discrete(name = "Involved") +
          sachin_theme_leg
   
fig3_un
ggsave("C:/Users/sachi/Box Sync/CTOP/Results for paper/fig_involvement_unadj 2020-03-20.png")


#### ADJUSTED ####

involved_adj <- involved  %>% filter( adjusted == "Adjusted")
fig3_adj <- ggplot(data = involved_adj, mapping = aes (x= involved, y = freq)) +
          geom_point(aes(color = EP), size = 3 , position = pd) +
          geom_errorbar(aes(ymin = LL, ymax = UL, width = 0.1, color = EP), position = pd) +
           scale_y_continuous(name = "Predicted likelihood", 
                             limits = c(0.0,0.11), breaks = c(seq(from = 0.0, to = 0.10, by = 0.02)),
                             labels = scales::percent_format(accuracy = 1)
                             ) +
          scale_x_discrete(name = "Involved") +
          sachin_theme_leg
fig3_adj
ggsave("C:/Users/sachi/Box Sync/CTOP/Results for paper/fig_involvement_adj 2020-03-20.png")
   


## ADJUSTED- BAR ##
involved_adj <- involved  %>% 
                  filter(adjusted == "Adjusted") %>%
                  mutate(sig = ifelse(involved == "Pharmacist", "N", "Y")) %>%
                  mutate(aster = ifelse(sig == "Y" & EP == "EP","*",NA))

involved_adj$involved <- as.factor(involved_adj$involved)
involved_adj$involved <- factor(involved_adj$involved, levels = c("Nonclinical health staff", "Clinician", "Pharmacist"))



#no differences were sig

fig_involvement_adj_bar <- ggplot(data = involved_adj, aes (x= involved, y = freq, fill = EP)) +
          geom_bar(stat = "identity", position = "dodge", width = .5) +
          scale_y_continuous(name = "Predicted likelihood", 
                             limits = c(0.0,0.10), breaks = c(seq(from = 0.0, to = 0.10, by = 0.02)),
                             labels = scales::percent_format(accuracy = 1)
                             ) +
          scale_x_discrete(name = "Involved", labels = wrap_format(10)) +
          sachin_theme_leg  
          



fig_involvement_adj_bar
ggsave("C:/Users/sachi/Box Sync/CTOP/Results for paper/fig_involvement_adj bar 2020-06-05.png", height = 4, width = 6)


```


### Figure 2: time to close
```{r}
library(readr)
library(tidyverse)

##clean up data
km <- read_csv("C:\\Users\\sachi\\Box Sync\\data\\cipher\\processed data\\ctop_apex_first_20200605.csv")

km <- filter(km, count_of_issues > 0) %>% 
      mutate(issue_closed = ifelse(max_close_time>14, 0, 1)) %>% 
      mutate(tt_close = ifelse(max_close_time>14, 14, max_close_time)) %>% 
      select(tt_close, issue_closed, EP, recode_race, ins, Disposition, Top3_service, Ethnicity, Marital_Status, age, LOS, Elixhauser)

km$EP <- as.factor(km$EP)



#run analysis
#ref: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html
library(survival)
library(survminer)
library(lubridate)
library(gtsummary)

#unadjusted
coxph(Surv(tt_close, issue_closed) ~ EP, data = km)

#adjusted
coxph(Surv(tt_close, issue_closed) ~ (EP + recode_race + ins + Disposition + Top3_service + Ethnicity + Marital_Status + age + LOS + Elixhauser), data = km) %>% gtsummary::tbl_regression(exp = TRUE) 

coxph(Surv(tt_close, issue_closed) ~ (EP + recode_race + ins + Disposition + Top3_service + Ethnicity + Marital_Status + age + LOS + Elixhauser), data = km)

fit <- coxph(Surv(tt_close, issue_closed) ~ (EP + recode_race + ins + Disposition + Top3_service + Ethnicity + Marital_Status + age + LOS + Elixhauser), data = km) 

fit <- coxph(Surv(tt_close, issue_closed) ~ (strata(EP2) +age), data = km) 

ggadjustedcurves(fit, data = km,  method = "marginal", reference = km)

ggadjustedcurves(res.cox, data = km)
surv_adjustedcurves(res.cox)
plot(res.cox)
fit <- survfit(res.cox, newdata = new_df)
ggsurvplot(fit, conf.int = TRUE, palette = "Dark2", 
           censor = FALSE, surv.median.line = "hv")

```


